### 2016-08-05 R/ITHIM.Rmd
First pass of NHTS data almost complete

DAYV2PUB <- within(DAYV2PUB, {
    URBRUR <- as.factor(URBRUR)
    HHSTFIPS <- as.factor(HHSTFIPS)
    MSACAT <- as.factor(MSACAT)
    TRPTRANS <- as.factor(TRPTRANS)
    })

@@

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(convertToAgeClass(R_AGE))
)
})

nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
sum(nMatrix$count) # 305649 number of people in the PERSON table

@@

NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))

foo <- lapply(split(NHTS, NHTS$SEX),function(x) split(x, x$AGE))
counts <- lapply(foo, function(x) lapply(x, countPeople))

>>>>

DAYV2PUB <- within(DAYV2PUB, {
    URBRUR <- as.factor(URBRUR)
    HHSTATE <- as.factor(HHSTATE)
    MSACAT <- as.factor(MSACAT)
    TRPTRANS <- as.factor(TRPTRANS)
    })

@@

convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

tripTime <- function(data, mode = "walk"){
 return(sum(subset(data, AT == mode)$TRVLCMIN, na.rm = TRUE))
}

reformatData <- function(data){
foo <- matrix(unlist(data), nrow = 7, ncol = 2, dimnames = list(paste0("ageClass",2:8), c("F","M")))
return(foo[,c("M","F")])
}

NHTS.person <- with(PERV2PUB,{
                    data.frame(HOUSEID = HOUSEID,
                               PERSONID = PERSONID,
                               SEX = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA))),
                               AGE = as.factor(convertToAgeClass(R_AGE)),
                               MSA = ifelse(MSASIZE %in% 1:5, TRUE, FALSE),
                               HHS = as.factor(convertToHHS(HHSTATE))

)
})

NHTS.person <- subset( NHTS.person, MSA == TRUE )

listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))

counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))

countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)

meanWalkList <- mapply(walkMatrix, countMatrix, FUN = "/", SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = "/", SIMPLIFY = FALSE)

```{r NHTSArchive, echo = FALSE, eval = FALSE}
nMatrix <- ddply( subset(NHTS.person, !is.na(AGE) & !is.na(SEX)), .(SEX, AGE), summarise, count = length(PERSONID))
sum(nMatrix$count) # 305649 number of people in the PERSON table

@@

NHTS %>% group_by(SEX, AGE) %>% dplyr::summarize(count = countPeople(data.frame(HOUSEID = HOUSEID, PERSONID = PERSONID)))
```
